---
title: "DecoupleR"
output: html_document
date: "2023-08-07"
---

For this script is used the full transcriptome data, not the SCT normalized data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, quiet=TRUE)
knitr::opts_knit$set(root.dir = '/home/rstudio/')
```

```{r, include=FALSE}
setwd('../rstudio/mod/')
# install.packages('Hmisc')
library(Hmisc)
# devtools::install_github("nicolash2/ggdendroplot")
# install.packages('reshape2')
# install.packages('ggplotify')
library(ggplotify)
library(reticulate)
library(remotes)

# reticulate::install_miniconda(update = TRUE, force = TRUE)
# reticulate::conda_install('r-reticulate', 'leidenalg')
# BiocManager::install('GSVA')
# BiocManager::install("decoupleR", ask = FALSE, force = TRUE)
# remotes::install_github('https://github.com/saezlab/OmnipathR.git')
# BiocManager::install(c("GenomeInfoDb", "XVector", "Biostrings", "KEGGREST", "S4Vectors", "IRanges", "annotate", "AnnotationDbi", "genefilter", "sva"))
# install.packages('devtools', quiet=TRUE)
# install.packages('Hmisc')
# devtools::install_github("data2intelligence/SpaCET", quiet=TRUE)
# install.packages('msigdbr')
# remotes::install_github('https://github.com/saezlab/decoupleR.git')
library(RColorBrewer)
library(Matrix)
library(magrittr)
library(dplyr)
library(ggplot2)
library(spdep)
library(kableExtra)
library(magrittr)
library(dplyr)
library(harmony)
library(decoupleR)
library(tibble)
library(tidyr)
library(factoextra)
library(viridis)
library(gridExtra)
library(stringr)
library(pheatmap)
library(msigdbr)
library(Seurat)
library(stats)
options(Seurat.object.assay.version = "v5")
options(future.globals.maxSize = 3000 * 1024^2)  # Set to 3 GB

source('/home/rstudio/mod/bin/utils.R', 
       local = TRUE
       )

filter_mt_rb_genes = function(ds){
  mt_rb_genes = grepl('^MT',rownames(ds),  ignore.case = TRUE) | grepl('^RPL',rownames(ds),  ignore.case = TRUE) | grepl('^RPS',rownames(ds),  ignore.case = TRUE)
  ds = ds[mt_rb_genes == FALSE, ]
  return(ds)
}
```

```{r slide batches}
cluster.pal = 'Paired' # cluster palette
sources = c('S1', 'S2', 'S3', 'S4')
slide.section = data.frame('slides'= c(117, 117, 121, 121, 121, 117, 117, 121, 110, 110, 110, 110),
  row.names = c('S1_C1', 'S1_D1', 'S2_A1', 'S2_B1', 'S2_C1','S3_A1' ,'S3_B1' ,'S3_D1' ,'S4_A1' ,'S4_B1' ,'S4_C1','S4_D1'))
```

```{r set a source/sample palette}
if(length(unique(sources)) == 1){
  brew.pal.sample = brewer.pal(n = ns, name = 'Set1')
} else {
  palettes=list()
  palettes[['S1']] = c('#e06666', '#ecb8b8')
  palettes[['S2']] = c('#28f618', '#25d283', '#23bfbb')
  palettes[['S3']] = c('#e2f4c7', '#eae374', '#f9d62e')
  palettes[['S4']] = c('#8d8bf1', '#5843a2', '#41296a', '#8638b7')

  brew.pal.sample = c()
  
  for(s in 1:length(unique(sources))){
    s.name = sources[s]
    brew.pal.sample = c(brew.pal.sample, palettes[[s.name]])
  }
get.brewer.pal = colorRampPalette(brewer.pal(16, cluster.pal))
}

ct_map_main = list(
  "Activated B cells" = "B/Plasma cells" ,
  "Astrocytes" = "Astrocytes",
  "CAFs" = "CAFs",
  "CD4+ T cells" = "T cells",
  "CD8+ T cells TCF7+" = "T cells",
  "CD8+ T cells TOX+" = "T cells",
  "DC3" = "Myeloid cells",
  "Endothelial cells" = "Endothelial cells",
  "MDM" = "Myeloid cells",
  "MDM FTL+" = "Myeloid cells",
  "Microglia"  = "Microglia",
  "Monocytes"     = "Myeloid cells",
  "NK cells"  = "NK cells",
  "Naïve B cells"  = "B/Plasma cells",
  "Neurons"  = "Neurons",
  "Oligodendrocytes" = "Oligodendrocytes",
  "Pericytes"    = "Pericytes",
  "Plasma cells"     = "B/Plasma cells",
  "T-cell doublets" = "doublets",
  "Tfh-like cells" = "T cells",
  "Tregs" =  "T cells",
  "Tumor cells" = "Tumor cells",
  "cDC1" = "Myeloid cells",
  "cDC2" = "Myeloid cells"
)
ct_main = unique(unlist(ct_map_main))
ct_main = ct_main[ct_main != 'doublets']
ct_main = gsub('Tumor cells', 'Malignant', ct_main)
```

```{r load seurat object, fig.height=20}
# file.name = paste0('./mod/data/transcriptome_analysis/', paste0(sources, collapse ='-'), '_pre_processed_slides_merged.rds')
# file.name.STACAS = paste0('/home/rstudio/mod/data/transcriptome_analysis/STACAS_integrated.rds')
# # se.STACAS = readRDS(file.name.STACAS)
# 
# se1 = readRDS(file.name)
# se = NormalizeData(se, assay = 'Spatial')
# se = add.spacet(se, force_interface = TRUE)
# se$sample_tissue = paste0(se$sample_id, '-', se$interface)
# se$source = gsub('S', 'P', se$source) #change coding of patients from "S_n" to "P_n" 

se = readRDS('./mod/data/decoupleR_se.rds')


# se@reductions$umap.STACAS = se.STACAS@reductions$umap
# se.STACAS@reductions$umap.nc = se@reductions$umap

n.clusters = length(unique(se$seurat_clusters))
ns = length(unique(se$sample_id))

SpatialDimPlot(se, group.by = c('interface'), ncol = 4, image.alpha = 0, pt.size.factor = 2, crop=FALSE)
SpatialFeaturePlot(se, features = c('Malignant'), ncol = 4, image.alpha = 0, pt.size.factor = 2, crop=FALSE)

pdf('/home/rstudio/mod/results/decoupler/info/interface.pdf',
  height = 8,
  width = 12
)
print(SpatialDimPlot(se, group.by = c('interface'), ncol = 4, image.alpha = 0, pt.size.factor = 2, crop=FALSE))
dev.off()

for (ct in rownames(se@reductions$SpaCET)){
  se[[ct]] = se@reductions$SpaCET[ct, ]
  pdf(paste0('/home/rstudio/mod/results/decoupler/info/', gsub('/', '-', ct),'.pdf'),
  height = 8,
  width = 12
)
  print(SpatialFeaturePlot(se, features = ct, ncol = 4, image.alpha = 0, pt.size.factor = 2, crop=FALSE))
  dev.off()
}

```
As the metastases tend to be very uniform, we perform a pseudobulking on Tumor, Interface and Stroma regions as identified by SpaCET

As the metastases tend to be very uniform, we perform a pseudobulking on Tumor, Interface and Stroma regions as identified by SpaCET
```{r melanoma scRNAseq, fig.height=10}
file = '/home/rstudio/mod/data/naive_melanoma/scCancer_reference.RDS'
if (!file.exists(file)){
X = read.csv('./mod/data/naive_melanoma/X.csv', row.names = 1)
meta = read.csv('./mod/data/naive_melanoma/metadata.csv', row.names = 1)
X = as.matrix(X)
gc()
scCancer_reference = CreateSeuratObject(counts = t(X), meta.data = meta)
scCancer_reference$source = scCancer_reference$orig.ident
scCancer_reference$interface = ifelse(scCancer_reference$cell_type_main == "Tumor cells", "Tumor", "Stroma")
scCancer_reference = NormalizeData(scCancer_reference)
scCancer_reference = ScaleData(scCancer_reference)
scCancer_reference = FindVariableFeatures(scCancer_reference)
scCancer_reference = RunPCA(scCancer_reference, npcs = 50)
scCancer_reference = RunUMAP(scCancer_reference, dims = 1:50, n.neighbors = 25)
saveRDS(scCancer_reference, file=file)
} else {
  scCancer_reference = readRDS(file)
}

CAFs_positions = which(scCancer_reference@meta.data[["cell_type_fine"]] == 'CAFs')
scCancer_reference$cell_type_main[CAFs_positions] = 'CAFs'
DimPlot(scCancer_reference, group.by = c('source', 'cell_type_main', 'interface'))

scCancer_reference.agg = AggregateExpression(scCancer_reference,return.seurat = TRUE, assays = 'RNA', group.by = c('interface', 'source'))
scCancer_reference.agg$patient = gsub('.+_', '', scCancer_reference.agg$orig.ident)
scCancer_reference.agg = FindVariableFeatures(scCancer_reference.agg)
scCancer_reference.agg = RunPCA(scCancer_reference.agg, npcs = 15)
scCancer_reference.agg = RunUMAP(scCancer_reference.agg, dims = 1:15, n.neighbors = 12)
DimPlot(scCancer_reference.agg, group.by = c('source', 'interface'))
```

```{r melanocytes, fig.height=10}
file = './mod/data/scrnaseq/melanocytes/melanocyte_ref.rds'
if (!file.exists(file)){
X = read.csv('./mod/data/scrnaseq/melanocytes/GSE151091_raw_matrix.csv.gz', row.names = 1)
meta = read.csv('./mod/data/scrnaseq/melanocytes/GSE151091_Metadata.csv.gz', row.names = 1)
common_cells = intersect(colnames(X), rownames(meta))
normal_reference = CreateSeuratObject(counts = X[, common_cells], meta.data =meta[common_cells,])
normal_reference$interface = 'Stroma'
normal_reference$source = normal_reference$patient

normal_reference = NormalizeData(normal_reference)
normal_reference = ScaleData(normal_reference)
normal_reference = FindVariableFeatures(normal_reference)
normal_reference = RunPCA(normal_reference, npcs = 50)
normal_reference = RunUMAP(normal_reference, dims = 1:50, n.neighbors = 25)
saveRDS(normal_reference, file = file)
} else {
  normal_reference = readRDS(file)
}
DimPlot(normal_reference, group.by = c('class_1', 'source', 'interface'))

normal_reference.agg = AggregateExpression(normal_reference,return.seurat = TRUE, assays = 'RNA', group.by = c('interface', 'source'))
normal_reference.agg$patient = gsub('.+_', '', normal_reference.agg$orig.ident)
normal_reference.agg = FindVariableFeatures(normal_reference.agg)
normal_reference.agg = RunPCA(normal_reference.agg, npcs = 15)
normal_reference.agg = RunUMAP(normal_reference.agg, dims = 1:15, n.neighbors = 12)
DimPlot(normal_reference.agg, group.by = c('source', 'interface'))
```
```{r merge, fig.height=8}
# union_genes = intersect(rownames(se@assays[["Spatial"]]), rownames(normal_reference@assays$RNA))
# union_genes = intersect(union_genes, rownames(scCancer_reference@assays$RNA))

normal.ref.agg = AggregateExpression(normal_reference, group.by = c("source", "interface"), assays = "RNA", return.seurat = TRUE)
normal.ref.agg$source = sapply(strsplit(colnames(normal.ref.agg), "_"), `[`, 1)
normal.ref.agg = filter_mt_rb_genes(normal.ref.agg)
normal.ref.agg = NormalizeData(normal.ref.agg)
normal.ref.agg$interface = 'Stroma'
normal.ref.agg$origin = 'control_ref'


normal.refct.agg = AggregateExpression(normal_reference, group.by = c("source", "interface", "class_1"), assays = "RNA", return.seurat = TRUE)
normal.refct.agg$source = sapply(strsplit(colnames(normal.refct.agg), "_"), `[`, 1)
normal.refct.agg$cell_type = sapply(strsplit(colnames(normal.refct.agg), "_"), `[`, 2)
normal.refct.agg = filter_mt_rb_genes(normal.refct.agg)
normal.refct.agg = NormalizeData(normal.refct.agg)
normal.refct.agg$interface = 'Stroma'
normal.refct.agg$origin = 'control_ref_ct'
normal.refct.agg$name = paste(normal.refct.agg$cell_type, normal.refct.agg$source)

scCancer.agg = AggregateExpression(scCancer_reference, group.by = c("source", "interface"), assays = "RNA", return.seurat = TRUE)
scCancer.agg$source = sapply(strsplit(colnames(scCancer.agg), "_"), `[`, 1)
scCancer.agg$interface = sapply(strsplit(colnames(scCancer.agg), "_"), `[`, 2)
scCancer.agg = filter_mt_rb_genes(scCancer.agg)
scCancer.agg = NormalizeData(scCancer.agg)
scCancer.agg$origin = 'cancer_ref'
scCancer.agg$name = paste(scCancer.agg$interface, scCancer.agg$source)

scCancerct.agg = AggregateExpression(scCancer_reference, group.by = c("source", "interface", "cell_type_main"), assays = "RNA", return.seurat = TRUE)
scCancerct.agg$source = sapply(strsplit(colnames(scCancerct.agg), "_"), `[`, 1)
scCancerct.agg$interface = sapply(strsplit(colnames(scCancerct.agg), "_"), `[`, 2)
scCancerct.agg$cell_type = scCancerct.agg$cell_type_main
scCancerct.agg = filter_mt_rb_genes(scCancerct.agg)
scCancerct.agg = NormalizeData(scCancerct.agg)
scCancerct.agg$origin = 'cancer_ref_ct'
scCancerct.agg$cell_type = scCancerct.agg$cell_type_main
scCancerct.agg$name = paste(scCancerct.agg$cell_type, scCancerct.agg$source)

se.agg = AggregateExpression(se, group.by = c("source", "interface"), assays = "Spatial", return.seurat = TRUE)
se.agg$source = sapply(strsplit(colnames(se.agg), "_"), `[`, 1)
se.agg$interface = sapply(strsplit(colnames(se.agg), "_"), `[`, 2)
se.agg@assays$RNA = se.agg@assays$Spatial
DefaultAssay(se.agg) = 'RNA'
se.agg = filter_mt_rb_genes(se.agg)
se.agg = NormalizeData(se.agg)
se.agg$origin = 'mmint'
se.agg$cell_type =se.agg$interface
se.agg$name = paste(se.agg$source, se.agg$interface)
de.markers = FindAllMarkers(se.agg, group.by = 'source', assay = 'RNA')
top10_se = de.markers %>%
  dplyr::group_by(cluster) %>%
  dplyr::top_n(wt = -p_val_adj, n = 10) %>%
  dplyr::top_n(wt = avg_log2FC, n = 10)



se.merged = merge(x = se.agg, y = c(normal.refct.agg, scCancerct.agg))
se.merged = JoinLayers(se.merged)
se.merged = ScaleData(se.merged)
se.merged = FindVariableFeatures(se.merged, nfeatures = 7500)
se.merged = RunPCA(se.merged, npcs = 50)
se.merged = RunUMAP(se.merged, dims = 1:50)

se.merged.int = se.merged
se.merged.int[["RNA"]] <- split(se.merged[["RNA"]], f = se.merged$origin)
se.merged.int <- NormalizeData(se.merged.int)
se.merged.int <- FindVariableFeatures(se.merged.int)
se.merged.int <- ScaleData(se.merged.int)
se.merged.int <- RunPCA(se.merged.int)

se.merged.int <- FindNeighbors(se.merged.int, dims = 1:30, reduction = "pca")
se.merged.int <- FindClusters(se.merged.int, resolution = 2, cluster.name = "unintegrated_clusters")
se.merged.int <- RunUMAP(se.merged.int, dims = 1:30, reduction = "pca", reduction.name = "umap.unintegrated")
DimPlot(se.merged.int, reduction = "umap.unintegrated", group.by = c("origin", "cell_type"))

se.merged.int <- IntegrateLayers(
  object = se.merged.int, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "integrated.harmony",
  verbose = TRUE
)

se.merged.int <- FindNeighbors(se.merged.int, reduction = "integrated.harmony", dims = 1:30)
se.merged.int <- FindClusters(se.merged.int, resolution = 2, cluster.name = "harmony_clusters")
se.merged.int <- RunUMAP(se.merged.int, reduction = "integrated.harmony", dims = 1:30, reduction.name = "umap.harmony")
p1 <- DimPlot(
  se.merged.int,
  reduction = "umap.harmony",
  group.by = c("origin", "cell_type", "source"),
  combine = TRUE, label.size = 2
)

p1

p1 <- FeaturePlot(
  se.merged.int,
  reduction = "umap.harmony", features = top10_se$gene,
  combine = TRUE, label.size = 2
)
p1
```
```{r, fig.height=25}
DotPlot(se.merged.int, assay = 'RNA', features = top10_se$gene, colors_37 <- c(
  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf",
  "#393b79", "#ff9896", "#98df8a", "#ffbb78", "#c5b0d5", "#8c6d31", "#e7ba52", "#17becf", "#9edae5", "#dbdb8d",
  "#f7b6d2", "#c49c94", "#aec7e8", "#ffbb78", "#98df8a", "#ff9896", "#c5b0d5", "#8c6d31", "#e7ba52", "#17becf",
  "#9edae5", "#dbdb8d", "#f7b6d2", "#c49c94", "#aec7e8", "#ffbb78", "#98df8a"
), dot.scale = 8, split.by = "source") +
    RotatedAxis()

DimPlot(se.merged, group.by = c('origin', 'cell_type')) 
DimPlot(se.merged, group.by = c('origin', 'cell_type'), reduction = 'pca') 
p1

```
As the metastases tend to be very uniform, we perform a pseudobulking on Tumor, Interface and Stroma regions as identified by SpaCET
```{r decoupleR functions}
# takes in net (source and target from the database) and returns the se adapted assay
run_decoupler = function(se,
                         net,
                         model = 'ulm',
                         .source = 'source',
                         .target = 'target',
                         .mor = 'weight',
                         assay = 'SCT',
                         store_as = 'progeny',
                         batch_kw = c(),
                         pvalue_filter = FALSE,
                         p_all = FALSE,
                         p_cutoff=0.05,
                         minsize = 5) {

  acts = c()
  assay_name = paste0(model, '_', store_as)
  print(assay_name)
  
  net = net %>%
    mutate(across(.source, str_replace_all, '_', '-'))
  net = na.omit(net)
  
  if (length(batch_kw) == 0) {
    se$batch = 'all'
    batch_kw = 'batch'
  }
  # print(unique(se[[batch_kw]]))
  DefaultAssay(se) = assay
  for (source in as.matrix(unique(se[[batch_kw]]))) {
    gc()
    print(paste('Processing patient sample:', source))
    se.partial = se[, se[[batch_kw]] == source]
    # print(se.partial)
    
    if (assay == 'Spatial') {
      mat = as.matrix(se.partial@assays[[assay]]$data)
    } else {
      mat = as.matrix(se.partial@assays[[assay]]$data)
    }
    # print(mat[1:5, 1:5])
    
    
    if (model == 'ulm'){
    if (.mor != '') {
      acts_source = run_ulm(
        mat = mat,
        net = net,
        .source = .source,
        .target = .target,
        .mor = .mor,
        minsize = minsize,
      )
    } else {
      acts_source = run_ulm(
        mat = mat,
        net = net,
        .source = .source,
        .target = .target,
        minsize = minsize,
      )
    }} else if (model == 'mlm'){
      if (.mor != '') {
      acts_source = run_mlm(
        mat = mat,
        net = net,
        .source = .source,
        .target = .target,
        .mor = .mor,
        minsize = minsize,
      )
      
      
      } 
      else {
      acts_source = run_mlm(
        mat = mat,
        net = net,
        .source = .source,
        .target = .target,
        minsize = minsize,
      )
    }
    } else if (model=='consensus') {
            acts_source = decouple(
        mat = mat,
        net = net,
        minsize = minsize,
        statistics = c('ulm', 'mlm')
        
      )
            acts_source = run_consensus(acts_source)
    } else {
      print('The model is not implemented yet')
    }
    
    acts = rbind(acts, acts_source)
  }
  
  acts$p_value_adj = p.adjust(acts$p_value, method = 'fdr')
  
  if (pvalue_filter == TRUE){ # modify it so it removes anything above 0.05
    if (p_all == FALSE){
      sig_acts = unique(acts[acts['p_value_adj'] <= p_cutoff, ][['source']])
    } else{
      sig_acts = setdiff(unique(acts[['source']]), unique(acts[acts['p_value_adj'] > p_cutoff, ][['source']]))
      if (length(sig_acts) == 0 ){
        print(sig_acts)
        print('No results')
        break
      }
    }
    acts = acts[acts[['source']] %in% sig_acts, ]
  }
  
  se[[assay_name]] = acts %>%
    filter(statistic == model) %>%
    pivot_wider(id_cols = 'source',
                names_from = 'condition',
                values_from = 'score') %>%
    column_to_rownames('source') %>%
    Seurat::CreateAssayObject(.)
  
  
  message(print(acts[1:10,]))
  p_vals = acts %>%
    filter(statistic == model) %>%
    pivot_wider(id_cols = 'source',
                names_from = 'condition',
                values_from = 'p_value_adj') %>%
    column_to_rownames('source')
  # print('pvals')
  # print(p_vals[1:5, 1:5])
  
  se[[assay_name]]@misc$pvals = p_vals[rownames(se[[assay_name]]),]
  
  

  
  se[[assay_name]]@misc$acts_long = acts
  print(se[[assay_name]])
  # Change assay
  DefaultAssay(object = se) = assay_name
  
  # Scale the data
  se = ScaleData(se)
  se@assays[[assay_name]]@data = se@assays[[assay_name]]@scale.data
  
  
  
  return(se)
}

QC_plot = function(se,
                   reduction = 'umap',
                   features = c(),
                   demo_slides = c('S1_C1', 'S2_B1', 'S3_D1', 'S4_A1'),
                   assay = 'ulm_progeny',
                   verbose = TRUE) {
  DefaultAssay(se) = assay
  if (length(features == 0)) {
    top_features = rownames(se)
  }
  
  max_features = 20
  if (length(features) > max_features) {
    print(paste(
      'Selecting top',
      max_features,
      'features out of',
      length(features)
    ))
    feature_ranges = apply(se@assays[[assay]]@data, 1, max) - apply(se@assays[[assay]]@data, 1, min)
    feature_ranges = feature_ranges[order(feature_ranges, decreasing = TRUE)]
    features = names(feature_ranges)[1:max_features]
    if (verbose == TRUE) {
      print(feature_ranges)
      print('Features with the largest span over all sources:')
      print(features)
    }
  } else {
    print(paste('features:', length(features)))
  }
  p1 <- DimPlot(
    se,
    group.by = 'interface',
    reduction = reduction,
    label = TRUE,
    pt.size = 0.25,
    alpha = 0.75,
    cols = c('darkgrey',  'black', 'gold1')
  ) + ggtitle('Tissue categories')
  
  # print('ffff')
  p2 = FeaturePlot(
    se,
    reduction = reduction,
    ncol = 3,
    features = features,
    combine = FALSE
  )
  p2 = lapply(
    X = p2,
    FUN = function(x)
      x + theme(plot.title = element_text(size = 5)) &
      scale_colour_gradient2(
        low = 'blue',
        mid = 'white',
        high = 'red'
      )
  )
  
  p2 = marrangeGrob(p2,
                    ncol = 4,
                    nrow = 4,
                    top = assay)
  
  p3 = DimPlot(
    se,
    reduction = reduction,
    group.by = 'source',
    pt.size = 0.25,
    alpha = 0.75
  )
  print(p1 - p3)
  print(p2)
  
  for (slide in demo_slides) {
    print(
      SpatialFeaturePlot(
        se,
        features = features,
        ncol = 4,
        images = slide,
        image.alpha = 0.5
      )
    )
  }
}


plot_results = function(assay = 'ulm_progeny',
                        folder = '/home/rstudio/mod/results/progeny/',
                        reduction = 'umap',
                        verbose = TRUE) {
  dir.create(folder,
             recursive = TRUE,
             showWarnings = FALSE)
  print('Plotting results')
  pathways = rownames(se@assays[[assay]]@data)
  n_paths = length(pathways)
  
  
  if (n_paths > 100) {
    print(paste('Selecting top 100 features out of', n_paths))
    pth_ranges = apply(se@assays[[assay]]@data, 1, max) - apply(se@assays[[assay]]@data, 1, min)
    pth_ranges = pth_ranges[order(pth_ranges, decreasing = TRUE)]
    pathways = names(pth_ranges)[1:100]
    n_paths = 100
    if (verbose == TRUE) {
      print(pth_ranges)
    }
  }
  
  ncol = 4
  nrow = ceiling(length(pathways) / ncol)
  
  i = 1
  plt.list = list()
  
  is = 1
  plt.list.s = list()
  
  for (path in pathways) {
    print(path)
    p = SpatialFeaturePlot(
      object = se,
      features = path,
      pt.size = 1.75,
      image.alpha = 0,
      ncol = 4,
      combine = FALSE
    )
    p = lapply(
      X = p,
      FUN = function(x)
        x + theme(plot.title = element_text(size = 5))
    )
    p = marrangeGrob(p,
                     ncol = 4,
                     nrow = 3,
                     top = assay)
    png(
      paste0(folder, path, '.png'),
      width = 350 * 4,
      height = 350 * 3,
      units = "px"
    )
    print(p)
    # print(marrangeGrob(plt.list, nrow = 1, ncol = 1))
    dev.off()
    
    pdf(paste0(folder, path, '.pdf'),
        height = 5 * ncol,
        width = 5 * nrow)
    print(p)
    # print(marrangeGrob(plt.list, nrow = 1, ncol = 1))
    dev.off()
    # is=1
    ps = (
      FeaturePlot(
        se,
        features = path,
        pt.size = 0.6,
        reduction = reduction,
        split.by = 'source',
        combine = TRUE,
        keep.scale = 'feature'
      ) &
        scale_colour_gradient2(
          low = 'blue',
          mid = 'white',
          high = 'red'
        )
    )
    
    for (k in 1:4) {
      ps[[k]] = ps[[k]] + theme_void()
      plt.list.s[[is]] = ps[[k]]
      is = is + 1
    }
    cowplot::plot_grid(plt.list.s[[1]] |
                         plt.list.s[[2]] | plt.list.s[[3]] | plt.list.s[[4]])
    
    
    p = (
      FeaturePlot(
        se,
        features = path,
        pt.size = 0.6,
        reduction = reduction
      ) &
        scale_colour_gradient2(
          low = 'blue',
          mid = 'white',
          high = 'red'
        )
    )
    
    plt.list[[i]] = p
    i = i + 1
  }
  png(
    paste0(folder, 'scores_umap_split.png'),
    width = 300 * 4,
    height = 300 * length(unique(se$sample_id)),
    units = "px"
  )
  print(marrangeGrob(
    plt.list.s,
    ncol = 4,
    nrow = n_paths,
    layout_matrix = matrix(
      seq_len(4 * n_paths),
      nrow = n_paths,
      ncol = 4,
      byrow = TRUE
    )
  ))
  dev.off()
  
  png(
    paste0(folder, 'scores_umap.png'),
    width = 600 * ncol,
    height = 600 * nrow,
    units = "px"
  )
  print(marrangeGrob(plt.list, nrow = nrow, ncol = ncol))
  dev.off()
  
  pdf(
    paste0(folder, 'scores_umap.pdf'),
    width = 30 * ncol,
    height = 30 * nrow
  )
  print(marrangeGrob(plt.list, nrow = nrow, ncol = ncol))
  dev.off()
}



plot_heatmap = function(se,
                        assay,
                        donor,
                        star = TRUE,
                        verbose = TRUE) {
  dir.create(output_folder,
             recursive = TRUE,
             showWarnings = FALSE)
  pathways = rownames(se@assays[[assay]]@data)
  max_path = 10000
  if (length(pathways) > max_path) {
    print(paste(
      'Selecting top',
      max_path,
      'features out of',
      length(pathways)
    ))
    pth_ranges = apply(se@assays[[assay]]@data, 1, max) - apply(se@assays[[assay]]@data, 1, min)
    pth_ranges = pth_ranges[order(pth_ranges, decreasing = TRUE)]
    pathways = names(pth_ranges)[1:max_path]
    n_paths = max_path
    if (verbose == TRUE) {
      # print(pth_ranges)
    }
  } else{
    pathways = pathways
  }
  
  
  se$source = gsub('-', '_', se$source) # because we have issues with dashes?
  print(se$source)
  df <- t(as.matrix(se@assays[[assay]]@data[pathways, ])) %>%
    as.data.frame() %>%
    mutate(tissue = se$interface) %>%
    mutate(donor = se$source) %>%
    pivot_longer(
      cols = !c(donor, tissue),
      names_to = "source",
      values_to = "score"
    ) %>%
    group_by(donor, tissue, source) %>%
    summarise(mean = mean(score))
  
  # Transform to wide matrix
  
  top_acts_mat = df %>%
    pivot_wider(
      id_cols = c('donor', 'tissue'),
      names_from = 'source',
      values_from = 'mean'
    )
  cnames = paste0(top_acts_mat$donor, '_', top_acts_mat$tissue)
  
  annotation.df = as.data.frame(top_acts_mat[, c('donor', 'tissue')])
  annotation.df[sapply(annotation.df, is.character)] = lapply(annotation.df[sapply(annotation.df, is.character)],
                                                              as.factor)
  
  top_acts_mat$donor = NULL
  top_acts_mat$tissue = NULL
  
  top_acts_mat = t(top_acts_mat)
  colnames(top_acts_mat) = cnames
  rownames(annotation.df) = colnames(top_acts_mat)
  
  ### Prepare the pvalues
  act_areas = unique(paste0(se$interface, '_',  se$source))
  act_areas = Idents(se)
  df_p = t(as.matrix(se@assays[[assay]]@misc$pvals[pathways, ]))[act_areas, ]
  df_p = df_p %>%
    as.data.frame() %>%
    mutate(tissue = se$interface) %>%
    mutate(donor = se$source) %>%
    pivot_longer(
      cols = !c(donor, tissue),
      names_to = "source",
      values_to = "score"
    ) %>%
    group_by(donor, tissue, source) %>%
    summarise(p_value = mean(score))
  
  df_p$stars = cut(
    df_p$p_value,
    breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
    label = c("***", "**", "*", "")
  )
  
  top_acts_mat_star <- df_p %>%
    pivot_wider(
      id_cols = c('donor', 'tissue'),
      names_from = 'source',
      values_from = 'stars'
    )
  top_acts_mat_star$donor = NULL
  top_acts_mat_star$tissue = NULL
  
  top_acts_mat_star = t(top_acts_mat_star)
  colnames(top_acts_mat_star) = cnames # and now you have the pvalues
  
  
  # Choose color palette
  palette_length = 100
  my_color = colorRampPalette(c("Darkblue", "white", "red"))(palette_length)
  
  my_breaks <-
    c(seq(-2, 0, length.out = ceiling(palette_length / 2) + 1),
      seq(0.05, 2, length.out = floor(palette_length / 2)))
  
  # create the color palette
  tissue = c(
    Stroma = "darkgrey",
    Interface = "black",
    Tumor = "gold1"
  )
  tissue = tissue[c(unique(se$interface))]
  annoCol = list(tissue = tissue)
  

  # print('this is the donor palette')
  # print(donor)
  # print('se.source')
  # print(c(unique(se$source)))
  donor = donor[c(unique(se$source))]
  annoCol = append(list(donor = donor), annoCol)
  # print('this is the donor list')
  # print(unique(annotation.df$donor))
  
  annoCol = append(list(tissue = annoCol$tissue[unique(annotation.df$tissue)]),
                   c(list(donor = annoCol$donor[unique(annotation.df$donor)])))
  
  # print(annotation.df[1:5, ])
  # print(annoCol)
  
  if (star == TRUE){
      pph = pheatmap(
    top_acts_mat,
    annotation_col = annotation.df,
    annotation_colors = annoCol,
    border_color = NA,
    color = my_color,
    breaks = my_breaks,
    fontsize = 7,
    display_numbers = top_acts_mat_star
    # cluster_cols = FALSE
  )
  } else {
      pph = pheatmap(
    top_acts_mat,
    annotation_col = annotation.df,
    annotation_colors = annoCol,
    border_color = NA,
    color = my_color,
    breaks = my_breaks,
    fontsize = 3,

  )
  }

  return(pph)
  
}

plot_heatmap_ct = function(se,
                        assay,
                        donor,
                        cell_type,
                        star = TRUE,
                        verbose = TRUE) {
  dir.create(output_folder,
             recursive = TRUE,
             showWarnings = FALSE)
  pathways = rownames(se@assays[[assay]]@data)
  max_path = 10000
  if (length(pathways) > max_path) {
    print(paste(
      'Selecting top',
      max_path,
      'features out of',
      length(pathways)
    ))
    pth_ranges = apply(se@assays[[assay]]@data, 1, max) - apply(se@assays[[assay]]@data, 1, min)
    pth_ranges = pth_ranges[order(pth_ranges, decreasing = TRUE)]
    pathways = names(pth_ranges)[1:max_path]
    n_paths = max_path
    if (verbose == TRUE) {
      # print(pth_ranges)
    }
  } else{
    pathways = pathways
  }
  
  
  se$source = gsub('-', '_', se$source)
  se$cell_type = gsub('-', '_', se$cell_type)
  names(cell_type) = gsub('-', '_', names(cell_type))
  donor = gsub('-', '_', donor)# because we have issues with dashes?
  # print(se$source)
  df <- t(as.matrix(se@assays[[assay]]@data[pathways, ])) %>%
    as.data.frame() %>%
    mutate(tissue = se$interface) %>%
    mutate(donor = se$source) %>%
    mutate(cell_type = se$cell_type) %>%
    pivot_longer(
      cols = !c(donor, tissue, cell_type),
      names_to = "source",
      values_to = "score"
    ) %>%
    group_by(donor, tissue, cell_type, source) %>%
    summarise(mean = mean(score))
  
  # Transform to wide matrix
  
  top_acts_mat = df %>%
    pivot_wider(
      id_cols = c('donor', 'tissue', 'cell_type'),
      names_from = 'source',
      values_from = 'mean'
    )
  cnames = paste0(top_acts_mat$donor, '_', top_acts_mat$tissue, '_', top_acts_mat$cell_type)
  
  annotation.df = as.data.frame(top_acts_mat[, c('donor', 'tissue', 'cell_type')])
  annotation.df[sapply(annotation.df, is.character)] = lapply(annotation.df[sapply(annotation.df, is.character)],
                                                              as.factor)
  
  top_acts_mat$donor = NULL
  top_acts_mat$tissue = NULL
  top_acts_mat$cell_type = NULL
  
  top_acts_mat = t(top_acts_mat)
  colnames(top_acts_mat) = cnames
  rownames(annotation.df) = colnames(top_acts_mat)
  
  ### Prepare the pvalues
  act_areas = unique(paste0(se$interface, '_',  se$source, '_', se$cell_type))
  act_areas = Idents(se)
  df_p = t(as.matrix(se@assays[[assay]]@misc$pvals[pathways, ]))[act_areas, ]
  df_p = df_p %>%
    as.data.frame() %>%
    mutate(tissue = se$interface) %>%
    mutate(donor = se$source) %>%
    mutate(cell_type = se$cell_type)  %>%
    pivot_longer(
      cols = !c(donor, tissue, cell_type),
      names_to = "source",
      values_to = "score"
    ) %>%
    group_by(donor, tissue, cell_type, source) %>%
    summarise(p_value = mean(score))
  
  df_p$stars = cut(
    df_p$p_value,
    breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
    label = c("***", "**", "*", "")
  )
  print(df_p$stars)
  top_acts_mat_star <- df_p %>%
    pivot_wider(
      id_cols = c('donor', 'tissue', 'cell_type'),
      names_from = 'source',
      values_from = 'stars'
    )
  top_acts_mat_star$donor = NULL
  top_acts_mat_star$tissue = NULL
  top_acts_mat_star$cell_type = NULL
  
  top_acts_mat_star = t(top_acts_mat_star)
  colnames(top_acts_mat_star) = cnames # and now you have the pvalues

  # Choose color palette
  palette_length = 100
  my_color = colorRampPalette(c("Darkblue", "white", "red"))(palette_length)
  
  my_breaks <-
    c(seq(-2, 0, length.out = ceiling(palette_length / 2) + 1),
      seq(0.05, 2, length.out = floor(palette_length / 2)))
  
  # create the color palette
  tissue = c(
    Stroma = "darkgrey",
    Interface = "black",
    Tumor = "gold1"
  )
  tissue = tissue[c(unique(se$interface))]
  donor = donor[c(unique(se$source))]
  cell_type = cell_type[c(unique(se$cell_type))]
  annoCol = list(tissue = tissue)
  annoCol = append(list(donor = donor), annoCol)
  annoCol = append(list(cell_type = cell_type), annoCol)

  # print('this is the donor palette')
  # print(donor)
  # print('se.source')
  # print(c(unique(se$source)))
  
  
  # print('this is the donor list')
  # print(unique(annotation.df$donor))
  

  
  annoCol = append(list(tissue = annoCol$tissue[unique(annotation.df$tissue)]),
                   c(list(donor = annoCol$donor[unique(annotation.df$donor)]), 
                     list(cell_type = annoCol$cell_type[unique(annotation.df$cell_type)])))
  
  print(annoCol)
  print(annotation.df)
  

  
  # print(annotation.df[1:5, ])
  # print(annoCol)
  
  if (star == TRUE){
      pph = pheatmap(
    top_acts_mat,
    annotation_col = annotation.df,
    annotation_colors = annoCol,
    border_color = NA,
    color = my_color,
    breaks = my_breaks,
    fontsize = 7,
    display_numbers = top_acts_mat_star
    # cluster_cols = FALSE
  )
  } else {
      pph = pheatmap(
    top_acts_mat,
    annotation_col = annotation.df,
    annotation_colors = annoCol,
    border_color = NA,
    color = my_color,
    breaks = my_breaks,
    fontsize = 3,

  )
  }

  return(pph)
  
}


comp_corr = function(se,
                     assay,
                     cutoff = 0.5,
                     split_by = 'source',
                     method = 'spearman',
                     output_folder = '/home/rstudio/mod/results/decoupler/correlations/') {
  dir.create(output_folder, showWarnings = FALSE)
  
  maxl = 0
  for (i in rownames(se@assays[[assay]]@data)){if (nchar(i)>maxl){maxl = nchar(i)}}
  print(maxl)
  
  for (var in unique(se[[split_by]][, ]))  {
    print(var)
    correlation_matrix = data.frame(matrix(nrow = 0, ncol = length(ct_main)))
    colnames(correlation_matrix) = ct_main
    pval_matrix = correlation_matrix
    
    se_part = se[, se[[split_by]] == var]
    
    # subselect the acts_long on the adjusted p_value
    Ppt = se_part@assays[[assay]]@misc[["acts_long"]]
    Ppt = Ppt[Ppt$condition%in%colnames(se_part), ]
    Ppt = Ppt[Ppt$p_value_adj < 0.05, ]
    pth_passed = table(Ppt$source) > (dim(se_part)[2]*cutoff)
    pth_passed = names(pth_passed[pth_passed == TRUE])
    print(table(Ppt$source))
    print(length(pth_passed))
    
    Apt = se_part@assays[[assay]]@data[pth_passed, ]
    Act = se_part@reductions$SpaCET[ct_main, colnames(se_part)]
    
    for (pth in pth_passed){
      print(pth)
      features = Ppt[Ppt$source==pth, ]$condition
      
      print(SpatialFeaturePlot(se[, features], features = c(pth)))

      Apt_pth = Apt[pth , features]
      result = rcorr(t(Act[, features]),
                     (Apt[pth , features]), 
                     type = method)
      correlation_matrix[pth, ] = result$r['y', ct_main]
      pval_matrix[pth, ] = result$P['y', ct_main] 
      
    }
    
    pval_stars <- apply(pval_matrix, c(1, 2), function(x) {
      if (is.na(x)) {
        return(NA) # Keep NA as it is
      } else if (x < 0.001) {
        return("***")
      } else if (x < 0.01) {
        return("**")
      } else if (x < 0.05) {
        return("*")
      } else {
        return("") # No significance
      }
    })
    # decoupleR vs ct

    s = pheatmap(correlation_matrix, display_numbers = pval_stars, fontsize = 12)
    width = 0.5 * length(ct_main) + 0.1*maxl
    height = 0.5 * length(pth_passed) + 0.1*maxl
    print(width)
    print(height)
    print('THIS WAS W AND HEIGHT')
    
    pdf(
      paste0(
        output_folder,
        assay,
        '_',
        var,
        '_corr_ct_assay_pheatmap_padj.pdf'
      ),
      height = 0.5 * length(rownames(Apt)) + 0.1*maxl,
      width = 0.5 * length(rownames(Act)) + 0.1*maxl)
    print(s)
    dev.off()

  }
}


pheatmap_merged = function(assay_name,
                           se.agg,
                           scCancerct.agg,
                           normal.refct.agg,
                           cluster_cols = TRUE,
                           cluster_rows = TRUE) {
  mmint_passed = grepl("Tumor|Stroma", colnames(se.agg))
  mat = se.agg@assays[[assay_name]]$data
  colnames(mat) = paste(se.agg$interface, se.agg$source)
  mat = mat[, mmint_passed]
  pval = se.agg@assays[[assay_name]]@misc$pvals[, mmint_passed]
  
  cancer_passed = grepl("Tumor cells|CAFs", colnames(scCancerct.agg))
  mat_add = scCancerct.agg@assays[[assay_name]]$data
  colnames(mat_add) = scCancerct.agg$name
  mat_add = mat_add[, cancer_passed]
  
  common = intersect(rownames(mat), rownames(mat_add))
  mat = cbind(mat[common, ], mat_add[common, sort(colnames(mat_add))])
  
  pval_add = scCancerct.agg@assays[[assay_name]]@misc$pvals
  colnames(pval_add) = scCancerct.agg$name
  pval_add = pval_add[, cancer_passed]
  pval = cbind(pval[common, ], pval_add[common, sort(colnames(pval_add))])
  
  normal_passed = grepl("mel", colnames(normal.refct.agg), ignore.case = TRUE)
  normal.refct.agg$name = gsub('mel', 'Melanocytes', normal.refct.agg$name)
  normal.refct.agg$name = gsub('cyc-Melanocytes',
                               'Cycling Melanocytes',
                               normal.refct.agg$name)
  mat_add = normal.refct.agg@assays[[assay_name]]$data
  colnames(mat_add) = normal.refct.agg$name
  mat_add = mat_add[, normal_passed]
  
  common = intersect(rownames(mat), rownames(mat_add))
  mat = cbind(mat[common, ], mat_add[common, sort(colnames(mat_add))])
  
  pval_add = normal.refct.agg@assays[[assay_name]]@misc$pvals
  colnames(pval_add) = normal.refct.agg$name
  pval_add = pval_add[, normal_passed]
  pval = cbind(pval[common, ], pval_add[common, sort(colnames(pval_add))])
  
  pval_stars = apply(pval, c(1, 2), function(x) {
    cut(
      x,
      breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
      labels = c("***", "**", "*", "")
    )
  })
  
  palette_length = 100
  my_color = colorRampPalette(c("darkblue", "white", "red"))(palette_length)
  
  my_breaks <-
    c(seq(-2, 0, length.out = ceiling(palette_length / 2) + 1),
      seq(0.05, 2, length.out = floor(palette_length / 2)))
  my_colour = list(
    sources = donor_palette[names(donor_palette) %in% anno.df$sources],
    cell_types = c(
      CAFs = "#6A3D9A",
      Cycling_Melanocytes = "palegreen4",
      Melanocytes = "palegreen",
      Stroma = 'darkgrey',
      Tumor = 'gold1',
      Tumor_cells = 'lemonchiffon'
    )
  )
  
  split_entry <- function(entry) {
    # Split at the last space
    parts = strsplit(entry, " ")[[1]]
    cell_type = paste(parts[-length(parts)], collapse = " ")
    sources = parts[length(parts)]
    return(list(cell_type = cell_type, sources = sources))
  }
  split_list = lapply(colnames(mat), split_entry)
  cell_types = sapply(split_list, function(x)
    x$cell_type)
  cell_types = gsub(' ', '_', cell_types)
  sources = sapply(split_list, function(x)
    x$source)
  sources = gsub('-', '_', sources)
  anno.df = data.frame(row.names = colnames(mat), cell_types, sources)
  
  pph = pheatmap(
    mat,
    cluster_cols = cluster_cols,
    cluster_rows = cluster_rows,
    display_numbers = pval_stars,
    color = my_color,
    fontsize = 12,
    breaks = my_breaks,
    annotation_col = anno.df,
    annotation_colors = my_colour
  )
  pph
  return(pph)
}

get_gene_pheatmap = function(se_list, genes, donor_palette){

  data = data.frame(row.names = genes)
  cell_types = c()
  sources = c()
  MITF = c()
  
  for (se in se_list){
  DefaultAssay(se) = 'RNA'
  se.part = se[genes, ] 
  cell_types = c(cell_types, as.character(se.part$cell_type))
  # MITF = c(MITF, se@assays$ulm_collecttri_TF_aggregate@data['MITF', ])
  sources = c(sources, se.part$source)
  genes = rownames(se.part)
  data_add = as.matrix(se.part@assays$RNA@layers$data)
  colnames(data_add) = colnames(se.part)
  rownames(data_add) = rownames(se.part)
  data = cbind(data[genes, ], data_add[genes, ])
  }
  
  # MITF = t(data.frame(MITF))
  # data = rbind( data, MITF)
  print(data)
  cell_types = gsub('mel', 'Melanocytes', cell_types)
  cell_types = gsub('cyc-Melanocytes', 'Cycling Melanocytes', cell_types)
  sources = gsub('-', '_', sources)
cell_types <- gsub("[/ -]", "_", cell_types)
colnames(data) = gsub('_', ' ', colnames(data))
print(table(cell_types))

  
  palette_length = 100
  my_color = colorRampPalette(c("darkblue", "white", "red"))(palette_length)
  cell_palette = c(
      CAFs = "#6A3D9A",
      Cycling_Melanocytes = "palegreen4",
      Melanocytes = "palegreen",
      Stroma = 'darkgrey',
      Tumor = 'gold1',
      Tumor_cells = 'lemonchiffon',
  cyc_krt = 'orangered3',      # Light salmon for cycling keratinocytes (more distinct)
  dendritic = '#1F78B4',    # Light blue for dendritic cells (closer to immune cells)
  eccrine = 'pink',      # Dark brown for eccrine (epithelial-related)
  krt = 'orangered2',          # OrangeRed for keratinocytes (distinct from melanocytes)
  mast = '#FF990F',         # Indian red for mast cells (myeloid-related)
  T_cell = 'blue',       # Lime green for T-cells
  B_Plasma_cells = '#4169E1', # Royal blue for B/Plasma cells
  CNS_cells = 'purple',    # Dark orchid for CNS cells
  Endothelial_cells = '#FF69B4', # Hot pink for endothelial cells
  Myeloid_cells = '#20B2AA', # Indian red for myeloid cells
  Stromal_cells = 'indianred', # Light sea green for stromal cells
  T_NK_cells = 'lightblue4'   # Forest green for T/NK cells
    )
  print(names(cell_palette))
  
    my_breaks <-
    c(seq(-2, 0, length.out = ceiling(palette_length / 2) + 1),
      seq(0.05, 2, length.out = floor(palette_length / 2)))
  my_colour = list(
    sources = donor_palette[names(donor_palette) %in% sources],
    cell_types = cell_palette[names(cell_palette)%in% cell_types]
  )
  anno.df = data.frame(row.names = colnames(data), cell_types, sources)
 pph = pheatmap(data, cluster_cols = TRUE, color = my_color, annotation_col = anno.df, 
    annotation_colors = my_colour)
  
  return(pph)
}

get_count_data = function(se, assay){mat = as.data.frame(se@assays[[assay]]@layers$counts); colnames(mat) = colnames(se@assays[[assay]]); rownames(mat) = rownames(se@assays[[assay]]); return(mat)}
```

```{r deseq, fig.height=10}
library(DESeq2)
library(sva)

normal.refct.agg.red = normal.refct.agg[, normal.refct.agg$cell_type%in%c('mel', 'cyc-mel', 'krt')]
scCancer.agg$cell_type = scCancer.agg$interface
scCancer.agg$cell_type = gsub('Tumor', 'Tumor cells', scCancer.agg$cell_type)
pass.cancer = names(table(scCancer.agg$source)[table(scCancer.agg$source) > 1]) 
scCancer.agg.red = scCancer.agg[, scCancer.agg$source%in%pass.cancer]

# Create a merged count matrix from the selected datasets
counts = merge(get_count_data(scCancerct.agg, assay='RNA'), get_count_data(se.agg, assay='Spatial'), by=0) 
row.names(counts) = counts$Row.names
counts$Row.names = NULL
counts = merge(counts, get_count_data(normal.refct.agg, assay='RNA'), by=0)
row.names(counts) = counts$Row.names
counts$Row.names = NULL
counts = as.matrix(counts)

# create the associated metadata
coldata = data.frame(Sample = colnames(counts))
coldata$dataset = c(scCancerct.agg$origin, se.agg$origin, normal.refct.agg$origin)
coldata$cell_type = c(scCancerct.agg$cell_type, se.agg$cell_type, normal.refct.agg$cell_type)
coldata$source = c(scCancerct.agg$source, se.agg$source, normal.refct.agg$source)
coldata$interface = c(scCancerct.agg$interface, se.agg$interface, normal.refct.agg$interface)

dds <- DESeqDataSetFromMatrix(countData = counts, colData = coldata, design = ~ 1)
dds <- estimateSizeFactors(dds)
norm_counts <- counts(dds, normalized = TRUE)


adjusted_counts <- ComBat_seq(counts, batch = coldata$dataset)

log_adj <- log2(adjusted_counts + 1)
pca <- prcomp(t(log_adj), scale. = TRUE)
ggplot(as.data.frame(pca$x), aes(PC1, PC2, color = coldata$cell_type, shape = coldata$dataset)) +
  geom_point(size = 3)

row.names(coldata) = coldata$Sample
dds.se = CreateSeuratObject(adjusted_counts, meta.data = coldata)
dds.se = NormalizeData(dds.se)
dds.se = ScaleData(dds.se)
dds.se = FindVariableFeatures(dds.se)
dds.se = RunPCA(dds.se)
dds.se = RunUMAP(dds.se, dims = 1:50)

# control plots
genes = c('PMEL', 'MLANA', 'DCT',  'MITF', 'AXL', 'ZEB1', 'TWIST1', 'TYR', 'TYRP1', 'MC1R')

DimPlot(dds.se, group.by = c('dataset', 'cell_type', 'interface'))
pph = get_gene_pheatmap(c(dds.se[, dds.se$cell_type%in%c('mel', 'cyc-mel', 'Tumor', 'Stroma', 'Tumor cells', 'CAFs')]), 
                        genes, 
                        donor_palette = donor_palette )
FeaturePlot(dds.se, features = genes)
DimPlot(dds.se, group.by = c('interface', 'cell_type'))
SpatialFeaturePlot(se, features = genes[1], ncol = 4, image.alpha = 0)
SpatialFeaturePlot(se, features = genes[2], ncol = 4, image.alpha = 0)
SpatialFeaturePlot(se, features = genes[3], ncol = 4, image.alpha = 0)
SpatialFeaturePlot(se, features = genes[4], ncol = 4, image.alpha = 0)

# save the resulting figure
pdf('./mod/results/decoupler/mitf_associated_genes.pdf', height = 4, width = 12)
print(pph)
dev.off()

```

```{r palettes}
  donor_palette = c(
    'P1' = 'darkorchid3',
    'P2' = 'yellow4',
    'P3' = 'pink3',
    'P4' = 'dodgerblue4',
  # First group (blue/green shades)
  'A1011L' = 'turquoise3',
  'A1012M' = 'cadetblue3',
  'A1014L' = 'steelblue3',
  'A1015LM' = 'aquamarine3',
  'A1016LM' = 'darkslategray3',
  'A1017LM' = 'paleturquoise3',
  'A1020LM' = 'lightseagreen',
  'A1021M' = 'deepskyblue3',
  'A1022M' = 'mediumturquoise',
  'A1025L' = 'darkturquoise',
  'A1026L' = 'powderblue',
  'A1033M' = 'skyblue3',
  'A1038LM' = 'lightskyblue3',
  'A1046M' = 'cornflowerblue',
  'FS030_LM' = 'darkseagreen3',
  'FS043_LM' = 'mediumaquamarine',
  # Second group (red/orange shades)
'MBM05_sn' = 'indianred1',
  'MBM06_sn' = 'firebrick3',
  'MBM07_sn' = 'tomato3',
  'MBM08_sn' = 'coral3',
  'MBM09_sn' = 'salmon3',
  'MBM10_sn' = 'darkorange3',
  'MBM11_sn' = 'sienna3',
  'MBM12_sn' = 'chocolate3',
  'MBM13_sn' = 'peru',
  'MBM14_sn' = 'darkgoldenrod3',
  'MBM15_sn' = 'tan3',
  'MBM16_sn' = 'rosybrown3',
  'MBM17_sn' = 'lightcoral',
  'MBM18_sn' = 'orangered3',
  'MBM19_sn' = 'brown3',
  'MBM20_sn' = 'maroon3',
  'MBM21_sn' = 'violetred3'
)

scRNA_ct_palette = c(
  'CAFs' = "#6A3D9A",
  'cyc_krt' = 'orangered3',      # Light salmon for cycling keratinocytes (more distinct)
  'cyc_mel' = 'brown4',      # Tomato for cycling melanocytes (closer to tumor)
  'dendritic' = '#1F78B4',    # Light blue for dendritic cells (closer to immune cells)
  'eccrine' = 'lightblue',      # Dark brown for eccrine (epithelial-related)
  'krt' = 'orangered2',          # OrangeRed for keratinocytes (distinct from melanocytes)
  'mast' = '#FF990F',         # Indian red for mast cells (myeloid-related)
  'mel' = 'brown',          # Dark orange for melanocytes (closer to tumor cells)
  'T-cell' = '#32CD32',       # Lime green for T-cells
  'B/Plasma cells' = '#4169E1', # Royal blue for B/Plasma cells
  'CNS cells' = 'blue',    # Dark orchid for CNS cells
  'Endothelial cells' = '#FF69B4', # Hot pink for endothelial cells
  'Myeloid cells' = '#CD5C5C', # Indian red for myeloid cells
  'Stromal cells' = '#20B2AA', # Light sea green for stromal cells
  'T/NK cells' = '#228B22',   # Forest green for T/NK cells
  'Tumor cells' = '#FFD700',
  'Stroma' = 'darkgrey',
  'Tumor' = 'gold1'
)

```

```{r progeny, fig.height=8, warning=FALSE, paged.print=TRUE}
net = get_progeny(organism = 'human', top = 100) # 100 genes from the human database

assay = 'progeny_aggregate'
model = 'ulm'
assay_name = paste0(model, '_', assay) # ulm is the default in all of these
output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')


# cancer ref cell type
scCancerct.agg = run_decoupler(
  se = scCancerct.agg,
  model = model,
  net = net,
  assay = 'RNA',
  store_as = assay,
  minsize = 5,
  pvalue_filter = FALSE,
  p_all = FALSE,
)
pph = plot_heatmap_ct(scCancerct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_ct_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# cancer  pseudobulk
scCancer.agg = run_decoupler(
  se = scCancer.agg,
  model = model,
  net = net,
  assay = 'RNA',
  store_as = assay,
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(scCancer.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref cell types
normal.refct.agg = run_decoupler(
  se = normal.refct.agg,
  model = model,
  net = net,
  assay = 'RNA',
  store_as = assay,
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap_ct(normal.refct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_ct_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref
normal.ref.agg = run_decoupler(
  se = normal.ref.agg,
  model = model,
  net = net,
  assay = 'RNA',
  store_as = assay,
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(normal.ref.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# mmint dataset
se.agg = run_decoupler(
  se = se.agg,
  model = model,
  net = net,
  assay = 'Spatial',
  store_as = assay,
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(se.agg[, se.agg$interface %in% c('Stroma', 'Tumor')],
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_mmint_scores_pheatmap_merged.pdf'
  ),
  height = 0.1* length(pph$tree_row$labels) + 2,
  width = 0.1* length(pph$tree_col$labels) + 3
)
print(pph)
dev.off()
```
```{r, fig.height=20}
pheatmap_merged(assay_name = assay_name, se.agg=se.agg, scCancerct.agg=scCancerct.agg, normal.refct.agg=normal.refct.agg)
DefaultAssay(se) = 'Spatial'
SpatialFeaturePlot(se, features = c('PMEL', 'DCT', 'TYR'), crop = FALSE, pt.size.factor = 2)
```

```{r, fig.height=10}
# mmint dataset spatial analysis
assay = 'progeny'
model = 'ulm'
assay_name = paste0(model, '_', assay)

output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')
dir.create(output_folder, showWarnings = FALSE)


se = run_decoupler(
  se = se,
  model = 'ulm',
  net = net,
  assay = 'Spatial',
  store_as = assay,
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
  batch_kw = 'sample_id'
)

paths = row.names(se@assays[[assay_name]]@data)

for (path in 1:length(paths)) {
  s = SpatialFeaturePlot(
    se,
    features = row.names(se@assays[[assay_name]]@data)[path],
    ncol = 4,
    pt.size.factor = 2.5,
    image.alpha = 0,
    crop = FALSE,
    combine = TRUE
  )
  
  pdf(
    paste0(
      '/home/rstudio/mod/results/decoupler/', assay,'/',
      assay_name,
      '_',
      paths[path],
      '_mmint_scores_spatial.pdf'
    ),
    height = 10,
    width = 8
  )
  print(s)
  dev.off()
}


for (path in 1:length(paths)) {
  p_val = se@assays[[assay_name]]@misc$pvals[path, ]<=0.05
  s = SpatialFeaturePlot(
    se[, p_val],
    features = row.names(se@assays[[assay_name]]@data)[path],
    ncol = 4,
    pt.size.factor = 2.5,
    image.alpha = 0,
    crop = FALSE,
    combine = TRUE
  )
  
  pdf(
    paste0(
      '/home/rstudio/mod/results/decoupler/', assay,'/',
      assay_name,
      '_',
      paths[path],
      '_mmint_scores_spatial_p005.pdf'
    ),
    height = 10,
    width = 8
  )
  print(s)
  dev.off()
}


comp_corr(
  se,
  assay = assay_name,
  split_by = 'source',
  method = 'pearson',
  cutoff = 0.5,
  output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay,'/correlations/')
)
```


```{r msigdb hallmark analysis, fig.height=15, fig.width=15}
# net = get_resource('MSigDB', organism = 'human')
# msigdb_colections = unique(net[, 'collection'])
# collection = 'hallmark'
# net = net[net[, 'collection'] == collection, ]

net = msigdbr(species = "human",
              category = "H")

net = distinct(net,
               diff = paste0(gs_name, human_gene_symbol),
               .keep_all = TRUE)

net$mor = 1
net$gs_name = gsub('_', ' ', net$gs_name)
net$gs_name = gsub('HALLMARK ', '', net$gs_name)


assay = 'msigdb_aggregate'
model = 'ulm'
assay_name = paste0(model, '_', assay)
output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')

# cancer ref cell type
scCancerct.agg = run_decoupler(
  se = scCancerct.agg,
  net = net, 
  model = model,
  .source='gs_name',
  .target='human_gene_symbol',
  .mor='',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap_ct(scCancerct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_ct_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# cancer  pseudobulk
scCancer.agg = run_decoupler(
  se = scCancer.agg,
  net = net,
  model = model,
  .source='gs_name',
  .target='human_gene_symbol',
  .mor='',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(scCancer.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref cell types
normal.refct.agg = run_decoupler(
  se = normal.refct.agg,
  net = net, 
  model = model,
  .source='gs_name',
  .target='human_gene_symbol',
  .mor='',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap_ct(normal.refct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_ct_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref
normal.ref.agg = run_decoupler(
  se = normal.ref.agg,
  net = net, 
  model = model,
  .source='gs_name',
  .target='human_gene_symbol',
  .mor='',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(normal.ref.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# mmint dataset
se.agg = run_decoupler(
  se = se.agg,
  net = net, 
  model = model,
  .source='gs_name',
  .target='human_gene_symbol',
  .mor='',
  assay = 'Spatial', 
  store_as = assay, 
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(se.agg[, se.agg$interface %in% c('Stroma', 'Tumor')],
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_mmint_scores_pheatmap_merged.pdf'
  ),
  height = 0.05* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

pheatmap_merged(assay_name = assay_name, se.agg=se.agg, scCancerct.agg=scCancerct.agg, normal.refct.agg=normal.refct.agg)
```
```{r, fig.height=10}
assay = 'msigdb'
assay_name = paste0(model, '_', assay)

output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')
dir.create(output_folder, showWarnings = FALSE)

# mmint dataset spatial analysis
se = run_decoupler(
  se = se,
  model = model,
  net = net,
  assay = 'Spatial',
  store_as = assay,
  .source='gs_name',
  .target='human_gene_symbol',
  .mor='',
  minsize = 5,
  pvalue_filter = TRUE,
  p_all = FALSE,
  batch_kw = 'sample_id'
)
paths = row.names(se@assays[[assay_name]]@data)

for (path in 1:length(paths)) {
  if (path %% 10 == 0 ){ print(paste0(path, '/', length(paths)))}
  s = SpatialFeaturePlot(
    se,
    features = row.names(se@assays[[assay_name]]@data)[path],
    ncol = 4,
    pt.size.factor = 2.5,
    image.alpha = 0,
    crop = FALSE,
    combine = TRUE
  )
  
  pdf(
    paste0(
      '/home/rstudio/mod/results/decoupler/', assay,'/',
      assay_name,
      '_',
      paths[path],
      '_mmint_scores_spatial.pdf'
    ),
    height = 20,
    width = 16
  )
  print(s)
  dev.off()
}

comp_corr(
  se,
  assay = assay_name,
  split_by = 'source',
  method = 'pearson',
  output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay,'/correlations/')
)
```

```{r TF pseudobulk, fig.height=25, fig.width=25}
net = get_collectri(organism = "human", split_complexes=FALSE)

assay = 'collecttri_TF_aggregate'
model = 'ulm'
assay_name = paste0(model, '_', assay) # ulm is the default in all of these
## reuse the mat from before (keep it consistent between methods)
output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')

# cancer ref cell type
scCancerct.agg = run_decoupler(
  se = scCancerct.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap_ct(scCancerct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_ct_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# cancer  pseudobulk
scCancer.agg = run_decoupler(
  se = scCancer.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(scCancer.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref cell types
normal.refct.agg = run_decoupler(
  se = normal.refct.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap_ct(normal.refct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_ct_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref
normal.ref.agg = run_decoupler(
  se = normal.ref.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = FALSE,
)
pph = plot_heatmap(normal.ref.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_scores_pheatmap_merged.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# mmint dataset
p_cutoff = 0.001
se.agg = run_decoupler(
  se = se.agg[, se.agg$interface %in% c('Stroma', 'Tumor')],
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'Spatial', 
  store_as = assay, 
  minsize = 5, 
  p_cutoff = p_cutoff,
  pvalue_filter = TRUE,
  p_all = TRUE,
)
pph = plot_heatmap(se.agg[, se.agg$interface %in% c('Stroma', 'Tumor')],
                   assay = assay_name, star = FALSE,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_mmint_scores_pheatmap_merged_pall_', p_cutoff,'_nostar.pdf'
  ),
  height = 0.02* length(pph$tree_row$labels) + 4,
  width = 0.001* length(pph$tree_col$labels) + 3
)
print(pph)
dev.off()
```
```{r merged heatmaps generation, fig.height=15}
output_folder = '/home/rstudio/mod/results/decoupler/'
# for (assay_name in c('ulm_progeny_aggregate', 'ulm_msigdb_aggregate', 'ulm_collecttri_TF_aggregate')){
for (assay_name in c('ulm_msigdb_aggregate')){

  pph_clust = pheatmap_merged(assay_name = assay_name, se.agg = se.agg, scCancerct.agg = scCancerct.agg, normal.refct.agg = normal.refct.agg)
  pph = pheatmap_merged(assay_name = assay_name, se.agg = se.agg, scCancerct.agg = scCancerct.agg, normal.refct.agg = normal.refct.agg, cluster_cols = FALSE)
  
    pdf(
  paste0(
    output_folder,
    assay_name,
    '_comparison_scores_pheatmap.pdf'
  ),
  height = 0.2* length(pph_clust$tree_row$labels) + 4,
  # height = 20,
  width = 0.3* length(pph_clust$tree_col$labels) + 4
)
print(pph)
dev.off()

pdf(
  paste0(
    output_folder,
    assay_name,
    '_comparison_scores_pheatmap_clustered_col.pdf'
  ),
  height = 0.2* length(pph_clust$tree_row$labels) + 4,
  width = 0.3* length(pph_clust$tree_col$labels) + 4
)
print(pph_clust)
dev.off()
}
```
```{r, fig.height=10}
# mmint dataset spatial analysis
assay = 'collecttri_TF'
model = 'ulm'
assay_name = paste0(model, '_', assay)
output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')

dir.create(output_folder, showWarnings = FALSE)

se = run_decoupler(
  se = se,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'Spatial', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = FALSE,
  batch_kw = 'sample_id'
)

paths = row.names(se@assays[[assay_name]]@data)[1:50]

for (path in 1:length(paths)) {
  if (path %% 10 == 0 ){ print(paste0(path, '/', length(paths)))}
  s = SpatialFeaturePlot(
    se,
    features = row.names(se@assays[[assay_name]]@data)[path],
    ncol = 4,
    pt.size.factor = 2.5,
    image.alpha = 0,
    crop = FALSE,
    combine = TRUE
  )
  
  pdf(
    paste0(
      '/home/rstudio/mod/results/decoupler/', assay,'/',
      assay_name,
      '_',
      paths[path],
      '_mmint_scores_spatial.pdf'
    ),
    height = 20,
    width = 16
  )
  print(s)
  dev.off()
}

comp_corr(
  se,
  assay = assay_name,
  split_by = 'source',
  method = 'pearson',cutoff = 0.9,
  output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay,'/correlations/')
)
```

```{r TF pseudobulk, fig.height=25, fig.width=25}
net = get_collectri(organism = "human", split_complexes=FALSE)

assay = 'collecttri_TF_aggregate'
model = 'ulm'
assay_name = paste0(model, '_', assay) # ulm is the default in all of these
## reuse the mat from before (keep it consistent between methods)
output_folder = paste0('/home/rstudio/mod/results/decoupler/', assay, '/')

mat = as.matrix(se.agg@assays[['Spatial']]$data)
# check_corr(net)

# cancer ref cell type
scCancerct.agg = run_decoupler(
  se = scCancerct.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = TRUE,
)
pph = plot_heatmap_ct(scCancerct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_ct_scores_pheatmap_merged_pall.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# cancer  pseudobulk
scCancer.agg = run_decoupler(
  se = scCancer.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = TRUE,
)
pph = plot_heatmap(scCancer.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_cancer_ref_scores_pheatmap_merged_pall.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref cell types
normal.refct.agg = run_decoupler(
  se = normal.refct.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = TRUE,
)
pph = plot_heatmap_ct(normal.refct.agg,
                   assay = assay_name,
                   cell_type = scRNA_ct_palette,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_ct_scores_pheatmap_merged_pall.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# normal ref
normal.ref.agg = run_decoupler(
  se = normal.ref.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'RNA', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = TRUE,
)
pph = plot_heatmap(normal.ref.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_normal_ref_scores_pheatmap_merged_pall.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()

# mmint dataset
se.agg = run_decoupler(
  se = se.agg,
  net = net, 
  model= model,
  .source='source',
  .target='target',
  .mor = 'mor',
  assay = 'Spatial', 
  store_as = assay, 
  minsize = 5, 
  pvalue_filter = TRUE,
  p_all = TRUE,
)
pph = plot_heatmap(se.agg,
                   assay = assay_name,
                   donor = donor_palette)
print(pph)

pdf(
  paste0(
    output_folder,
    assay_name,
    '_mmint_scores_pheatmap_merged_pall.pdf'
  ),
  height = 0.2* length(pph$tree_row$labels) + 4,
  width = 0.2* length(pph$tree_col$labels) + 4
)
print(pph)
dev.off()
```
